---
title: "4.2_multiome_transplantations_analysis"
author: "Anna Schwager"
date: today
format: gfm
---


```{r}
#| label: silence
#| include: false

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Python setup

```{r}
#| label: reticulate

library(reticulate)
use_condaenv(
  "snap_py310",
  conda = "/Users/annaschwager/miniforge3/bin/conda",
  required = TRUE
)

#py_config()
```

```{python}
#| label: check-python

import sys
import numpy as np
import pandas as pd
from pathlib import Path
import scanpy as sc
import snapatac2 as snap
import skmisc

#print("sys.executable:", sys.executable)
#print("scanpy:", sc.__version__)
#print("snapatac2:", snap.__version__)

```

# R setup

```{r}
#| label: source

here::i_am("scripts/4.2_multiome_transplantations_analysis.qmd")
mainDir = here::here()
source(knitr::purl(file.path(mainDir,"scripts/global_variables.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir, "scripts/functions.Rmd"), quiet=TRUE))

set.seed(123)
```

# Setting directories and input files

```{r}
#| label: set_inputs

inputDir_local = file.path(outputDir, "4.1_multiome_transplantations_preprocessing", "objects")

outputDir_local = file.path(outputDir, "4.2_multiome_transplantations_analysis") ; if(!file.exists(outputDir_local)){dir.create(outputDir_local)}
outputDir_objects = file.path(outputDir_local, "objects") ; if(!file.exists(outputDir_objects)){dir.create(outputDir_objects)}
outputDir_plots = file.path(outputDir_local, "plots") ; if(!file.exists(outputDir_plots)){dir.create(outputDir_plots)}
outputDir_tables = file.path(outputDir_local, "tables") ; if(!file.exists(outputDir_tables)){dir.create(outputDir_tables)}

snap_dir <- file.path(outputDir_objects, "snap_export") ; if(!file.exists(snap_dir )){dir.create(snap_dir)}
```

# Load annotation

```{r}
#| label: annot

ref_inhouse <- read.delim(file.path(annotDir, "Landragin2025_CreN_mammary_gland_markers.txt"))
ref_visvader <- read.delim(file.path(annotDir, "Visvader2010_mammary_gland_markers.txt"))
```

# Load object

```{r read_rds3, message=F, warning=F}
#| label: load0

seurat <- readRDS(file.path(inputDir_local, "seurat_step3.rds"))
seurat <- readRDS(file.path(inputDir_local, "seurat_all.rds"))
```

# Exporting data from Seurat to SnapATAC2

## RNA
```{r}
#| label: snap1
#| eval: false

seurat <- JoinLayers(seurat, assay = "RNA_fl")
rna_assay <- GetAssayData(seurat, assay = "RNA_fl", layer = "counts")

Matrix::writeMM(rna_assay, file.path(snap_dir, "matrix_rna_fl.mtx"))
write.table(colnames(rna_assay), file.path(snap_dir, "barcodes_rna_fl.tsv"),
            quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)

genes_df <- data.frame(
  name  = rownames(rna_assay),
  stringsAsFactors = FALSE
)
write.table(genes_df, file.path(snap_dir, "features_rna_fl.tsv"),
            quote = FALSE, sep = "\t", row.names = FALSE, col.names = TRUE)

```

## H3K4me1
```{r}
#| label: snap2
#| eval: false

k4_assay <- GetAssayData(seurat, assay = "bin_5k_h3k4me1", layer = "counts")

Matrix::writeMM(k4_assay, file.path(snap_dir, "matrix_k4.mtx"))
write.table(colnames(k4_assay), file.path(snap_dir, "barcodes_k4.tsv"),
            quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)


gr <- StringToGRanges(rownames(k4_assay), sep = c(":", "-"))
bins_df <- data.frame(
  chrom = as.character(seqnames(gr)),
  start = start(gr),
  end   = end(gr),
  name  = rownames(k4_assay),
  stringsAsFactors = FALSE
)
write.table(bins_df, file.path(snap_dir, "features_k4.tsv"),
            quote = FALSE, sep = "\t", row.names = FALSE, col.names = TRUE)

```

# Spectral embeddings generation
## Full length RNA
```{python}
#| label: snap3
#| eval: false

snap_dir = Path(r.snap_dir) 
X = sc.read_mtx(snap_dir / "matrix_rna_fl.mtx").X.tocsr().transpose()
barcodes = pd.read_csv(snap_dir / "barcodes_rna_fl.tsv",
                       header=None, sep="\t")[0].values
feat = pd.read_csv(snap_dir / "features_rna_fl.tsv", sep="\t")

adata_rna_fl = sc.AnnData(X=X)
adata_rna_fl.obs_names = barcodes
adata_rna_fl.var_names = feat["name"].values
adata_rna_fl.X = adata_rna_fl.X.tocsr().astype(np.float64)

adata_rna_fl
```

```{python}
#| label: snap4
#| eval: false

sc.pp.highly_variable_genes(adata_rna_fl, flavor='seurat_v3', n_top_genes=3000)
adata_rna_fl = adata_rna_fl[:, adata_rna_fl.var.highly_variable]

sc.pp.normalize_total(adata_rna_fl, target_sum=1e4)
sc.pp.log1p(adata_rna_fl)

adata_rna_fl
```

```{python}
#| label: snap5
#| eval: false
 
snap.tl.spectral(adata_rna_fl, features=None)
```

## H3K4me1
```{python}
#| label: snap6
#| eval: false

snap_dir = Path(r.snap_dir) 
X = sc.read_mtx(snap_dir / "matrix_k4.mtx").X.tocsr().transpose()  
barcodes = pd.read_csv(snap_dir / "barcodes_k4.tsv", header=None, sep="\t")[0].values
feat = pd.read_csv(snap_dir / "features_k4.tsv", sep="\t")

adata_k4 = sc.AnnData(X=X)
adata_k4.obs_names = barcodes
adata_k4.var_names = feat["name"].values
adata_k4.var["chrom"] = feat["chrom"].values
adata_k4.var["start"] = feat["start"].values
adata_k4.var["end"]   = feat["end"].values

adata_k4.X = adata_k4.X.tocsr().astype(np.float64)
adata_k4
```

```{python}
#| label: snap7
#| eval: false

snap.pp.select_features(adata_k4, n_features = 300000)
snap.tl.spectral(adata_k4, features = 'selected', distance_metric='cosine')
```

# Import embeddings to Seurat
```{r}
#| label: snap8
#| eval: false

emb_rna_fl <- as.matrix(py$adata_rna_fl$obsm[["X_spectral"]])
rownames(emb_rna_fl) <- py$adata_rna_fl$obs_names$tolist()

emb_k4 <- as.matrix(py$adata_k4$obsm[["X_spectral"]])
rownames(emb_k4) <- py$adata_k4$obs_names$tolist()

```

```{r}
#| label: snap9
#| eval: false

seurat[["spectral_rna_fl"]] <- CreateDimReducObject(
  embeddings = emb_rna_fl,
  key       = "specrnafl_",
  assay     = "RNA_fl" 
)

seurat[["spectral_k4"]] <- CreateDimReducObject(
  embeddings = emb_k4,
  key       = "speck4_",
  assay     = "bin_5k_h3k4me1"
)

```

# Reload step 2
```{r}
#| label: save_step2
#| eval: false

saveRDS(seurat, file.path(outputDir_objects, "seurat_step2.rds"))

```

```{r}
#| label: load_step2

seurat <- readRDS(file.path(outputDir_objects, "seurat_step2.rds"))

```


# H3K4me1 clustering and annotation 
```{r}
#| label: umap_k4
#| eval: false

seurat <- RunUMAP(seurat,
                  assay = "bin_5k_h3k4me1",
                  seed.use = 42,
                  dims = 1:length(colnames(seurat@reductions[["spectral_k4"]])),
                  reduction = "spectral_k4",
                  reduction.name = "UMAP_spectral_k4")

seurat <- FindNeighbors(seurat,
                        reduction = "spectral_k4",
                        cluster.name = "spectral_k4_clusters",
                        graph.name = "spectral_k4_snn",
                        dims = 1:length(colnames(seurat@reductions[["spectral_k4"]])),
                        k.param = 10)

seurat <- FindClusters(seurat,
                       resolution = 0.1,
                       cluster.name = "spectral_k4_clusters",
                       graph.name = "spectral_k4_snn",
                       algorithm = 4)


DimPlot(seurat, cols = mypal, 
        reduction = "UMAP_spectral_k4",
        group.by = c("Exp", "facs_phenotype","annot_rna", "spectral_k4_clusters"),
        combine = TRUE)

FeaturePlot(seurat, 
        reduction = "UMAP_spectral_k4",
        features = c("frequency_count", "FRiP"),
        combine = TRUE)
```


## Batch correction
```{r}
#| label: batch_k4
#| eval: false

seurat <- RunHarmony(
  object        = seurat,
  group.by.vars = c("Exp", "frequency_count", "FRiP"),
  reduction.use = "spectral_k4",
  reduction.save = "spectral_k4_harmony",
  project.dim    = FALSE,
  plot_convergence = TRUE
)
  
```

```{r}
#| label: post_batch_umap
#| eval: false

seurat <- RunUMAP(seurat,
                  assay = "bin_5k_h3k4me1",
                  seed.use = 42,
                  metric = "cosine",
                  n.neighbors = 50,
                  min.dist = 0.6,
                  spread = 1.5,
                  dims = 1:length(colnames(seurat@reductions[["spectral_k4"]])),
                  reduction = "spectral_k4_harmony",
                  reduction.name = "UMAP_spectral_k4_harmony")

seurat <- FindNeighbors(seurat,
                        reduction = "spectral_k4_harmony",
                        cluster.name = "spectral_k4_harmony_clusters",
                        graph.name = "spectral_k4_harmony_snn",
                        dims = 1:length(colnames(seurat@reductions[["spectral_k4"]])),
                        k.param = 10)

seurat <- FindClusters(seurat,
                       graph.name   = "spectral_k4_harmony_snn",
                       resolution = 0.2,
                       cluster.name = "spectral_k4_harmony_clusters",
                       algorithm = 4)

```


```{r}
#| label: post_batch_umap2
#| eval: false

DimPlot(seurat, cols = mypal, 
        reduction = "UMAP_spectral_k4_harmony",
        group.by = c("Exp", "facs_phenotype",
                     "spectral_k4_harmony_clusters"),
        combine = TRUE)

FeaturePlot(seurat, 
        reduction = "UMAP_spectral_k4_harmony",
        features = c("log10_frequency_count", "FRiP"),
        combine = TRUE)
```


## Cell type scores
```{r}
#| label: scores
#| eval: false

seurat <- SCTransform(seurat, assay = "RNA_fl", new.assay.name = "SCT_fl", vars.to.regress = c("nCount_RNA_fl", "nFeature_RNA_fl", "Exp"))

```

Landragin
```{r}
#| label: scores2
#| eval: false

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_inhouse,
                 cell_type = "Basal",
                 assay = "SCT_fl",
                 score_name = "basal_rna_score_Landragin")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_inhouse,
                 cell_type = "Luminal H-S",
                 assay = "SCT_fl",
                 score_name = "hs_rna_score_Landragin")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_inhouse,
                 cell_type = "LP",
                 assay = "SCT_fl",
                 score_name = "lp_rna_score_Landragin")


seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_inhouse,
                 cell_type = "Basal",
                 assay = "ga_h3k4me1",
                 score_name = "basal_k4_score_Landragin")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_inhouse,
                 cell_type = "Luminal H-S",
                 assay = "ga_h3k4me1",
                 score_name = "hs_k4_score_Landragin")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_inhouse,
                 cell_type = "LP",
                 assay = "ga_h3k4me1",
                 score_name = "lp_k4_score_Landragin")

```

Visvader
```{r}
#| label: scores3
#| eval: false

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_visvader,
                 cell_type = "Basal",
                 assay = "SCT_fl",
                 score_name = "basal_rna_score_Visvader")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_visvader,
                 cell_type = "Luminal H-S",
                 assay = "SCT_fl",
                 score_name = "hs_rna_score_Visvader")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_visvader,
                 cell_type = "LP",
                 assay = "SCT_fl",
                 score_name = "lp_rna_score_Visvader")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_visvader,
                 cell_type = "Basal",
                 assay = "ga_h3k4me1",
                 score_name = "basal_k4_score_Visvader")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_visvader,
                 cell_type = "Luminal H-S",
                 assay = "ga_h3k4me1",
                 score_name = "hs_k4_score_Visvader")

seurat <- AddCellTypeScore(seurat_obj = seurat,
                 marker_table = ref_visvader,
                 cell_type = "LP",
                 assay = "ga_h3k4me1",
                 score_name = "lp_k4_score_Visvader")

```

```{r}
#| label: scores4
#| eval: false

VlnPlot(seurat, c("basal_k4_score_Landragin", "lp_k4_score_Landragin", "hs_k4_score_Landragin"),
                   group.by = c("spectral_k4_harmony_clusters"),
                   pt.size = 0.5, cols = mypal, y.max = 5) +
                   stat_summary(fun.y = median, geom='point', size = 2,
                                colour = "black", position = position_dodge(width = 0.9))

VlnPlot(seurat, c("basal_k4_score_Visvader", "lp_k4_score_Visvader", "hs_k4_score_Visvader"),
                   group.by = c("spectral_k4_harmony_clusters"),
                   pt.size = 0.5, cols = mypal, y.max = 5) +
                   stat_summary(fun.y = median, geom='point', size = 2,
                                colour = "black", position = position_dodge(width = 0.9))

VlnPlot(seurat, c("basal_rna_score_Landragin", "lp_rna_score_Landragin", "hs_rna_score_Landragin"),
                   group.by = c("spectral_k4_harmony_clusters"),
                   pt.size = 0.5, cols = mypal) +
                   stat_summary(fun.y = median, geom='point', size = 2,
                                colour = "black", position = position_dodge(width = 0.9))

VlnPlot(seurat, c("basal_rna_score_Visvader", "lp_rna_score_Visvader", "hs_rna_score_Visvader"),
                   group.by = c("spectral_k4_harmony_clusters"),
                   pt.size = 0.5, cols = mypal) +
                   stat_summary(fun.y = median, geom='point', size = 2,
                                colour = "black", position = position_dodge(width = 0.9))

```

0 - basal
2 - LP2
3 - LP1
4 - HS
5 - ? 
6 - ? 
7 - ?


## Cluster annotation

```{r}
#| label: annot1
#| eval: false

DefaultAssay(seurat) <- "SCT_fl"
PrepSCTFindMarkers(seurat, assay = "SCT_fl")
Idents(seurat) <- "spectral_k4_harmony_clusters"

c5_markers <- FindMarkers(seurat, ident.1 = "5")
c6_markers <- FindMarkers(seurat, ident.1 = "6")
c7_markers <- FindMarkers(seurat, ident.1 = "7")

```

5 - endothelial cells
6 - peripheral glia
7 - unknown epithelial


#  Reload step 3

```{r}
#| label: save_step3
#| eval: false

saveRDS(seurat, file.path(outputDir_objects, "seurat_step3.rds"))

```

```{r}
#| label: load_step3

seurat <- readRDS(file.path(outputDir_objects, "seurat_step3.rds"))

```


# Remove non-epithelial clusters 

```{r}
#| label: remove
#| eval: true

seurat <- subset(
  seurat,
  subset = !(spectral_k4_harmony_clusters %in% c("5", "6"))
)
```


# K4 Diffusion map

```{r}
#| label: diffusion1
#| eval: false

X <- Seurat::Embeddings(seurat[["spectral_k4_harmony"]])[, 1:16]
dm <- DiffusionMap(X)

dc <- destiny::eigenvectors(dm)
dc <- as.matrix(dc)

seurat[["diffusion_spectral_k4"]] <- Seurat::CreateDimReducObject(
  embeddings = dc,
  key = "DIF_",
  assay = "bin_5k_h3k4me1"
)
```



```{r}
#| label: diffusion2
#| eval: false

DimPlot(seurat, reduction = "diffusion_spectral_k4", group.by = c("facs_phenotype"),
        cols = mypal)

DimPlot(seurat, reduction = "diffusion_spectral_k4", group.by = c("facs_phenotype"), dims = c(1,3),
        cols = mypal)


```

# Global pseudotime K4

```{r}
#| label: pseudotime_global
#| eval: false


dpt <- DPT(dm)
saveRDS(dpt, file.path(outputDir_objects, "DPT.rds"))
```


```{r}
#| label: pseudotime_global2
#| eval: true

dpt <- readRDS(file.path(outputDir_objects, "DPT.rds"))

```


```{r}
#| label: pseudotime_global3
#| eval: true

plot(dpt, col_by = 'DPT1')
plot(dpt, root = 1, paths_to = c(2, 3), col_by = 'branch')

```

# Branch annotation
```{r}
#| label: branch_annot1
#| eval: false

FeaturePlot(seurat,
            reduction = "diffusion_spectral_k4",
            features = c("basal_rna_score_Landragin")) +
  xlim(-0.05, 0.05) +
  ylim(-0.05, 0.05)

FeaturePlot(seurat,
            reduction = "diffusion_spectral_k4",
            features = c("lp_rna_score_Landragin")) +
  xlim(-0.05, 0.05) +
  ylim(-0.05, 0.05)

FeaturePlot(seurat,
            reduction = "diffusion_spectral_k4",
            features = c("hs_rna_score_Landragin")) +
  xlim(-0.05, 0.05) +
  ylim(-0.05, 0.05)

```
1 = HS
2 = basal
3 = LP


```{r}
#| label: branch_annot2
#| eval: false

# extract branch annotation
branch <- dpt$branch
branch[is.na(branch)] <- "ambiguous"
names(branch) <- rownames(destiny::eigenvectors(dpt@dm))

branch_label <- as.character(branch)

branch_label[branch_label == "1"] <- "HS"
branch_label[branch_label == "2"] <- "basal"
branch_label[branch_label == "3"] <- "LP"
branch_label[is.na(branch_label)] <- "ambiguous"

branch_label <- factor(
  branch_label,
  levels = c("basal", "HS", "LP", "ambiguous")
)

# extract tip annotation
tips_mat <- dpt@tips
tips_mat[is.na(tips_mat)] <- FALSE

is_tip <- apply(tips_mat, 1, any)
names(is_tip) <- rownames(tips_mat)


seurat <- Seurat::AddMetaData(
  seurat,
  metadata = branch,
  col.name = "DPT_branch"
)

seurat <- Seurat::AddMetaData(
  seurat,
  metadata = branch_label,
  col.name = "DPT_branch_label"
)

seurat <- Seurat::AddMetaData(
  seurat,
  metadata = is_tip,
  col.name = "DPT_is_tip"
)
```


## Annotating branch-ambiguous cells by RNA and FACS

Defining RNA score thresholds 
```{r}
#| label: score_thresholds
#| eval: false

VlnPlot(seurat, cols = mypal,
        features = c("basal_rna_score_Landragin", "lp_rna_score_Landragin", "hs_rna_score_Landragin"),
        group.by = "DPT_branch_label")

```

1.5 for all scores


```{r}
#| label: missing_annot
#| eval: false

md <- seurat@meta.data

md$annot_k4_rna_facs <- as.character(md$DPT_branch_label)

is_amb <- md$DPT_branch_label == "ambiguous"

basal_only <- is_amb &
  md$basal_rna_score_Landragin > 1.5 &
  md$lp_rna_score_Landragin    < 1.5 &
  md$hs_rna_score_Landragin    < 1.5

lp_only <- is_amb &
  md$lp_rna_score_Landragin    > 1.5 &
  md$basal_rna_score_Landragin < 1.5 &
  md$hs_rna_score_Landragin    < 1.5

hs_only <- is_amb &
  md$hs_rna_score_Landragin    > 1.5 &
  md$basal_rna_score_Landragin < 1.5 &
  md$lp_rna_score_Landragin    < 1.5

md$annot_k4_rna_facs[basal_only] <- "basal"
md$annot_k4_rna_facs[lp_only]    <- "LP"
md$annot_k4_rna_facs[hs_only]    <-  "HS"

still_amb <- is_amb & md$annot_k4_rna_facs == "ambiguous" & md$facs_phenotype != "all_epith"
md$annot_k4_rna_facs[still_amb] <- as.character(md$facs_phenotype[still_amb])

seurat@meta.data <- md
table(seurat@meta.data$annot_k4_rna_facs)

```


# Basal -> LP pseudotime, K4

```{r}
#| label: pt_basal_lp
#| eval: false

s <- subset(
  seurat,
  subset = !(DPT_branch_label %in% c("HS"))
)

X_BLP <- Seurat::Embeddings(s[["spectral_k4_harmony"]])[, 1:16]
dm_BLP <- DiffusionMap(X_BLP)

dc_BLP <- destiny::eigenvectors(dm_BLP)
dc_BLP <- as.matrix(dc_BLP)

s[["diffusion_spectral_k4_BLP"]] <- Seurat::CreateDimReducObject(
  embeddings = dc_BLP,
  key = "DIFBLP_",
  assay = "bin_5k_h3k4me1"
)

```

```{r}
#| label: pt_basal_lp2
#| eval: false

dpt_BLP  <- DPT(dm_BLP)
```


```{r}
#| label: pt_basal_lp3
#| eval: false

plot(dpt_BLP, col_by = 'DPT1')
```

Add DPT and metadata to the object
```{r}
#| label: pt_basal_lp4
#| eval: false

dpt_global <- dpt$DPT1
names(dpt_global) <- rownames(destiny::eigenvectors(dm))

dpt_basal_lp <- dpt_BLP$DPT1
names(dpt_basal_lp) <- rownames(destiny::eigenvectors(dm_BLP))


seurat <- Seurat::AddMetaData(
  seurat,
  metadata = dpt_global,
  col.name = "DPT_global"
)

seurat <- Seurat::AddMetaData(
  seurat,
  metadata = dpt_basal_lp,
  col.name = "DPT_basal_lp"
)

```

#  Reload step 4
```{r}
#| label: save_step4
#| eval: false

saveRDS(seurat, file.path(outputDir_objects, "seurat_step4.rds"))

```

```{r}
#| label: load_step4

seurat <- readRDS(file.path(outputDir_objects, "seurat_step4.rds"))

```



# RNA Diffusion map 

```{r}
#| label: diffusion_rna_batch
#| eval: false

seurat <- RunHarmony(
  object        = seurat,
  group.by.vars = c("Exp"),
  reduction.use = "spectral_rna_fl",
  reduction.save = "spectral_rna_fl_harmony",
  project.dim    = FALSE,
  plot_convergence = TRUE
)
  
```


```{r}
#| label: diffusion_rna1
#| eval: false

X <- Seurat::Embeddings(seurat[["spectral_rna_fl_harmony"]])[, 1:18]
dmr <- DiffusionMap(X)

dcr <- destiny::eigenvectors(dmr)
dcr <- as.matrix(dcr)

seurat[["diffusion_rna_fl"]] <- Seurat::CreateDimReducObject(
  embeddings = dcr,
  key = "DIFrna_",
  assay = "RNA_fl"
)

```


```{r}
#| label: diffusion_rna2
#| eval: false

DimPlot(seurat, reduction = "diffusion_rna_fl", group.by = c("facs_phenotype"),
        cols = cols)

DimPlot(seurat, reduction = "diffusion_rna_fl", group.by = c("DPT_branch_label"), 
        cols = cols)

```

Remove 3 extreme outliers (likely non-epithelial cells)
```{r}
#| label: diffusion_rna3
#| eval: false

emb <- Embeddings(seurat[["diffusion_rna_fl"]])
seurat$DIFrna_3 <- emb[, 3]

rna <- GetAssayData(seurat, assay="RNA_fl", layer="counts")
rna_cells <- colnames(rna)

pass_rna <- intersect(
  WhichCells(seurat, expression = DIFrna_3 > -0.1),
  rna_cells
)

rna_pass <- rna[, pass_rna, drop = FALSE]

seurat[["RNA_fl"]] <- CreateAssayObject(counts = rna_pass)

```

Recalculate harmony
```{r}
#| label: diffusion_rna4
#| eval: false

seurat <- RunHarmony(
  object        = seurat,
  group.by.vars = c("Exp"),
  reduction.use = "spectral_rna_fl",
  reduction.save = "spectral_rna_fl_harmony",
  project.dim    = FALSE,
  plot_convergence = TRUE
)
  
```


```{r}
#| label: diffusion_rna5
#| eval: false

X <- Seurat::Embeddings(seurat[["spectral_rna_fl_harmony"]])[, 1:18]
dmr <- DiffusionMap(X)

dcr <- destiny::eigenvectors(dmr)
dcr <- as.matrix(dcr)

seurat[["diffusion_rna_fl"]] <- Seurat::CreateDimReducObject(
  embeddings = dcr,
  key = "DIFrna_",
  assay = "RNA_fl"
)

```



```{r}
#| label: diffusion_rna6
#| eval: false

DimPlot(seurat, reduction = "diffusion_rna_fl", group.by = c("facs_phenotype"),
        cols = cols)

DimPlot(seurat, reduction = "diffusion_rna_fl", group.by = c("DPT_branch_label"))

```


# Pseudotime RNA
## Global 

```{r}
#| label: pt_rna1
#| eval: false

dptr <- DPT(dmr)

```

```{r}
#| label: pt_rna2
#| eval: false

plot(dptr, col_by = 'DPT1')
#plot(dptr, root = 1, paths_to = c(2, 3), col_by = 'branch')

```

## Basal -> LP

```{r}
#| label: pt_rna3
#| eval: false


s <- subset(
  seurat,
  subset = !(DPT_branch_label %in% c("HS"))
)

X_BLP_rna <- Seurat::Embeddings(s[["spectral_rna_fl_harmony"]])[, 1:18]
dm_BLP_rna <- DiffusionMap(X_BLP_rna)

dc_BLP_rna <- destiny::eigenvectors(dm_BLP_rna)
dc_BLP_rna <- as.matrix(dc_BLP_rna)

s[["diffusion_spectral_rna_BLP"]] <- Seurat::CreateDimReducObject(
  embeddings = dc_BLP_rna,
  key = "DIFBLPrna_",
  assay = "RNA_fl"
)

```

```{r}
#| label: pt_rna4
#| eval: false

dpt_BLP_rna  <- DPT(dm_BLP_rna)
```


```{r}
#| label: pt_rna5
#| eval: false

plot(dpt_BLP_rna, col_by = 'DPT1')
plot(dpt_BLP_rna, root = 1, paths_to = c(2, 3), col_by = 'branch')
```


Add DPT and metadata to the object
```{r}
#| label: pt_rna6
#| eval: false

# extract numerical values for the dominant diffusion pseudotime
dpt_global_rna <- dptr$DPT1
names(dpt_global_rna) <- rownames(destiny::eigenvectors(dmr))

dpt_basal_lp_rna <- dpt_BLP_rna$DPT1
names(dpt_basal_lp_rna) <- rownames(destiny::eigenvectors(dm_BLP_rna))


seurat <- Seurat::AddMetaData(
  seurat,
  metadata = dpt_global_rna,
  col.name = "DPT_global_rna"
)

seurat <- Seurat::AddMetaData(
  seurat,
  metadata = dpt_basal_lp_rna,
  col.name = "DPT_basal_lp_rna"
)

```


#  Reload step 5

```{r}
#| label: save_step5
#| eval: false

saveRDS(seurat, file.path(outputDir_objects, "seurat_step5.rds"))

```


```{r}
#| label: load_step5

seurat <- readRDS(file.path(outputDir_objects, "seurat_step5.rds"))

```


Remove "hybrid" facs cells as there are only three of them
```{r}
#| label: sub

seurat <- subset(
  seurat,
  subset = !(facs_phenotype %in% c("hybrid"))
)

```


```{r}
#| label: days

seurat$day <- NA_character_
cells_exp67 <- seurat$Exp %in% c("exp6", "exp7")

seurat$day[cells_exp67 & seurat$facs_phenotype == "all_epith"] <- "d4.5"
seurat$day[cells_exp67 & seurat$facs_phenotype != "all_epith"] <- "d0"

seurat$day_facs <- NA_character_
seurat$day_facs[seurat$day == "d0" & seurat$facs_phenotype == "basal"] <- "d0_basal"
seurat$day_facs[seurat$day == "d0" & seurat$facs_phenotype == "luminal"] <- "d0_luminal"
seurat$day_facs[seurat$day == "d4.5" & seurat$facs_phenotype == "all_epith"] <- "d4.5_all"

```


# Plots 
```{r}
#| label: color

cols <- c(
   basal   = "#E64B35",
   HS      = "#754f75",
   LP      = "#ba8cba",
   luminal = "#e0d0e0",
   ambiguous = "grey",
   d0_basal = "#E64B35",
   d0_luminal = "#ba8cba",
   d4.5_all = "#91D1C2E5",
   all_epith = "#91D1C2E5",
   transitioning = "#F39B7FE5"
 )

```


## Donuts
### FACS 
Data from FlowJo

```{r}
#| label: d4.5_facs

# predefined composition from FlowJo
df_pct <- tibble::tibble(
  annot_k4 = c("basal", "luminal", "int"),
  frac = c(0.23, 0.38, 0.39)
) |>
  mutate(
    n = frac * 100,                             # use 100 as total to match "percent"
    label = paste0(annot_k4, "\n(", round(frac*100, 1), "%)")
  )

sum_n <- sum(df_pct$n)

circle_outer <- data.frame(x = 2.5, y = seq(0, sum_n, length.out = 400))
circle_inner <- data.frame(x = 1.5, y = seq(0, sum_n, length.out = 400))

p_donut_facs_d4.5 <- ggplot(df_pct, aes(x = 2, y = n, fill = annot_k4)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  scale_fill_manual(values = cols, drop = FALSE) +
  theme_void() +
  theme(legend.title = element_blank()) +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  geom_path(
    data = circle_outer, aes(x = x, y = y),
    inherit.aes = FALSE, color = "black", linewidth = 0.8
  ) +
  geom_path(
    data = circle_inner, aes(x = x, y = y),
    inherit.aes = FALSE, color = "black", linewidth = 0.8
  )

p_donut_facs_d4.5

```

```{r}
#| label: d4.5_facs_save
#| eval: false

ggsave(file.path(outputDir_plots, "donut_facs_d0.pdf"),
       plot = p_donut_facs_d0,
       device = "pdf",
       units = "mm",
       width = 90,
       height = 90)

ggsave(file.path(outputDir_plots, "donut_facs_d4.5.pdf"),
       plot = p_donut_facs_d4.5,
       device = "pdf",
       units = "mm",
       width = 90,
       height = 90)
```


### H3K4me1 + RNA + FACS
```{r}
#| label: d4.5_omics

df_d45 <- seurat@meta.data |>
  dplyr::filter(day == "d4.5", !is.na(annot_k4_rna_facs)) |>
  dplyr::mutate(annot_k4_rna_facs = as.character(annot_k4_rna_facs)) |>
  dplyr::count(annot_k4_rna_facs, name = "n") |>
  dplyr::mutate(
    frac  = n / sum(n),
    label = paste0(annot_k4_rna_facs, "\n(", n, ")")
  )

sum_n <- sum(df_d45$n)

circle_outer <- data.frame(
  x = 2.5,
  y = seq(0, sum_n, length.out = 400)
)

circle_inner <- data.frame(
  x = 1.5,
  y = seq(0, sum_n, length.out = 400)
)

p_donut_d4.5 <- ggplot(df_d45, aes(x = 2, y = n, fill = annot_k4_rna_facs)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  scale_fill_manual(values = cols, drop = FALSE) +
  theme_void() +
  theme(legend.title = element_blank()) +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3) +
  # outer outline
  geom_path(
    data = circle_outer,
    aes(x = x, y = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = 0.8
  ) +
  # inner outline (hole)
  geom_path(
    data = circle_inner,
    aes(x = x, y = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = 0.8
  )

p_donut_d4.5

```


```{r}
#| label: d0_omics

df_d0 <- seurat@meta.data |>
  dplyr::filter(day == "d0", !is.na(annot_k4_rna_facs)) |>
  dplyr::mutate(annot_k4_rna_facs = as.character(annot_k4_rna_facs)) |>
  dplyr::count(annot_k4_rna_facs, name = "n") |>
  dplyr::mutate(
    frac  = n / sum(n),
    label = paste0(annot_k4_rna_facs, "\n(", n, ")")
  )

sum_n <- sum(df_d0$n)

circle_outer <- data.frame(
  x = 2.5,
  y = seq(0, sum_n, length.out = 400)
)

circle_inner <- data.frame(
  x = 1.5,
  y = seq(0, sum_n, length.out = 400)
)

p_donut_d0 <- ggplot(df_d0, aes(x = 2, y = n, fill = annot_k4_rna_facs)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  scale_fill_manual(values = cols, drop = FALSE) +
  theme_void() +
  theme(legend.title = element_blank()) +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3) +
  # outer outline
  geom_path(
    data = circle_outer,
    aes(x = x, y = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = 0.8
  ) +
  # inner outline (hole)
  geom_path(
    data = circle_inner,
    aes(x = x, y = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = 0.8
  )

p_donut_d0

```


```{r}
ggsave(file.path(outputDir_plots, "joint_annot_donut_d0.pdf"),
       plot = p_donut_d0,
       device = "pdf",
       units = "mm",
       width = 90,
       height = 90)

ggsave(file.path(outputDir_plots, "joint_annot_donut_d4.5.pdf"),
       plot = p_donut_d4.5,
       device = "pdf",
       units = "mm",
       width = 90,
       height = 90)
```


## Diffusion by FACS
```{r}
#| label: diffusion_map_prep

DefaultAssay(seurat) <- "ga_h3k4me1"  # TO keep Cells(seurat) consistent

emb <- Seurat::Embeddings(seurat, reduction = "diffusion_spectral_k4")[, 1:3, drop = FALSE]
df <- data.frame(
  cell = rownames(emb),
  DC1  = emb[, 1],
  DC2  = emb[, 2],
  DC3  = emb[, 3]
)

df <- cbind(
  df,
  seurat@meta.data[df$cell, , drop = FALSE]
)

```


```{r}
#| label: diffusion_map

df_rest <- df[df$facs_phenotype != "all_epith", ]
df_trans <- df[df$facs_phenotype == "all_epith", ]

p_12_f_d0 <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = facs_phenotype),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "facs_phenotype",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 colored by facs_phenotype",
          x = "DC1",
          y = "DC2"
        )

p_12_f_d0


p_12_f_all <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = facs_phenotype),
          size = 1.8,
          alpha = 1
        ) +
        geom_point(
          data = df_trans,
          aes(DC1, DC2, color = facs_phenotype),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "facs_phenotype",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 colored by facs_phenotype",
          x = "DC1",
          y = "DC2"
        )

p_12_f_all

```

```{r}
#| label: diffusion_map_save
#| eval: false

ggsave(file.path(outputDir_plots, "diff_map_facs_d0.pdf"),
       plot = p_12_f_d0,
       device = "pdf",
       units = "mm",
       width = 115,
       height =100)

ggsave(file.path(outputDir_plots, "diff_map_facs_all.pdf"),
       plot = p_12_f_all,
       device = "pdf",
       units = "mm",
       width = 115,
       height =100)

```



## Diffusion by H3K4me1 branch

```{r}
#| label: diffusion_branch

df_rest <- df[df$facs_phenotype != "all_epith", ]
df_trans <- df[df$facs_phenotype == "all_epith", ]

p_12_branch_d0 <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = DPT_branch_label),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "DPT_branch_label",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 colored by annot_k4",
          x = "DC1",
          y = "DC2"
        )
p_12_branch_d0

p_12_branch_all <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = DPT_branch_label),
          size = 1.8,
          alpha = 1
        ) +
        geom_point(
          data = df_trans,
          aes(DC1, DC2, color = DPT_branch_label),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "DPT_branch_label",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 colored by annot_k4",
          x = "DC1",
          y = "DC2"
        )
p_12_branch_all

```

```{r}
#| label: diffusion_branch_save
#| eval: false

ggsave(file.path(outputDir_plots, "diff_map_branch_d0.pdf"),
       plot = p_12_branch_d0,
       device = "pdf",
       units = "mm",
       width = 120,
       height =100)

ggsave(file.path(outputDir_plots, "diff_map_branch_all.pdf"),
       plot = p_12_branch_all,
       device = "pdf",
       units = "mm",
       width = 120,
       height =100)

```


## Diffusion by joint annot

```{r}
#| label: diffusion_joint_annot

df_rest <- df[df$facs_phenotype != "all_epith", ]
df_trans <- df[df$facs_phenotype == "all_epith", ]
df_basal <- df[df$facs_phenotype != "all_epith" & df$annot_k4_rna_facs == "basal", ]

p_12_branch_d0 <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = annot_k4_rna_facs),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "annot_k4_rna_facs",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 colored by joint annot",
          x = "DC1",
          y = "DC2"
        )
p_12_branch_d0

p_12_branch_all <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = annot_k4_rna_facs),
          size = 1.8,
          alpha = 1
        ) +
        geom_point(
          data = df_trans,
          aes(DC1, DC2, color = annot_k4_rna_facs),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "annot_k4_rna_facs",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 colored by annot_k4_rna_facs",
          x = "DC1",
          y = "DC2"
        )
p_12_branch_all


```

```{r}
#| label: diffusion_joint_annot_save
#| eval: false

ggsave(file.path(outputDir_plots, "diff_map_branch_d0_joint_annot.pdf"),
       plot = p_12_branch_d0,
       device = "pdf",
       units = "mm",
       width = 120,
       height =100)

ggsave(file.path(outputDir_plots, "diff_map_branch_all_joint_annot.pdf"),
       plot = p_12_branch_all,
       device = "pdf",
       units = "mm",
       width = 120,
       height =100)

ggsave(file.path(outputDir_plots, "diff_map_basal.pdf"),
       plot = p_only_basal,
       device = "pdf",
       units = "mm",
       width = 80,
       height =100)

```




## Diffusion, repopulated cells overlayed
```{r}
#| label: outline

library(sf)

pts <- st_as_sf(df_trans, coords = c("DC1", "DC2"), crs = NA)
r <- 0.0022

shape   <- st_union(st_buffer(pts, dist = r))
outline <- st_boundary(shape)

p_overlay <- ggplot() +
  geom_point(data = df_rest, aes(DC1, DC2),
             color = "grey80", size = 2.5) +
  geom_point(data = df_trans, aes(DC1, DC2),
             color = "#91D1C2E5", size = 2.5) +
  geom_sf(data = outline, inherit.aes = FALSE,
          color = "black", linewidth = 0.7) +
  theme_classic() +
  labs(x = "DC1", y = "DC2") +
  coord_sf(expand = FALSE) +
  theme(aspect.ratio = 1 / 0.78)

p_overlay
```

```{r}
#| label: outline_save
#| eval: false

ggsave(file.path(outputDir_plots, "diff_map_overlay_brown.pdf"),
       plot = p_overlay,
       device = "pdf",
       units = "mm",
       width = 140,
       height =100)

```


## RNA DPT
```{r}
#| label: plot_dpt_rna

df_pt <- df[!is.na(df$DPT_basal_lp_rna), ]

p_dpt_rna <- ggplot(df, aes(DC1, DC2)) +
                    geom_point(color = "grey80", size = 1.8) +
                    geom_point(
                      data = df_pt,
                      aes(color = DPT_basal_lp_rna),
                      size = 1.9
                    ) +
                    scale_color_viridis_c(
                      option = "magma",
                      name = "DPT_basal_lp_rna",
                      na.value = "grey80",
                      limits = c(min(df$DPT_basal_lp_rna, na.rm = TRUE), 2.5),
                      oob = scales::squish
                    ) +
                    theme_classic() +
                    labs(x = "DC1", y = "DC2")

p_dpt_rna

```

```{r}
#| label: save_dpt_rna
#| eval: false

ggsave(file.path(outputDir_plots, "dpt_rna.pdf"),
       plot = p_dpt_rna,
       device = "pdf",
       units = "mm",
       width = 140,
       height =100)

```

## K4 DPT

```{r}
#| label: plot_dpt_k4


df_pt <- df[!is.na(df$DPT_basal_lp), ]

p_dpt <- ggplot(df, aes(DC1, DC2)) +
                    geom_point(color = "grey80", size = 1.8) +
                    geom_point(
                      data = df_pt,
                      aes(color = DPT_basal_lp),
                      size = 1.9
                    ) +
                    scale_color_viridis_c(
                      option = "magma",
                      name = "DPT_basal_lp",
                      na.value = "grey80",
                    ) +
                    theme_classic() +
                    labs(x = "DC1", y = "DC2")
p_dpt

```

```{r}
#| label: save_dpt_k4
#| eval: false


ggsave(file.path(outputDir_plots, "dpt_k4.pdf"),
       plot = p_dpt,
       device = "pdf",
       units = "mm",
       width = 130,
       height =100)
```

## Ridge plots
### Global DPT
```{r}
#| label: plot_dpt_global
#| eval: false

p_gl_facs <- RidgePlot(seurat, features = c("DPT_global"), group.by = "facs_phenotype", cols = cols, sort = TRUE)
p_gl_branch <- RidgePlot(seurat, features = c("DPT_global"), group.by = "DPT_branch_label", cols = cols, sort = TRUE)

p_gl_facs
p_gl_branch

```


```{r}
#| label: plot_dpt_global2
#| eval: false

cells_use <- WhichCells(seurat, expression = !is.na(DPT_global_rna) & is.finite(DPT_global_rna))

p_gl_facs_rna <-RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_global_rna",
  group.by = "facs_phenotype",
  cols = cols,
  sort = TRUE
)

p_gl_branch_rna <-RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_global_rna",
  group.by = "DPT_branch_label",
  cols = cols,
  sort = TRUE
)

p_gl_facs_rna
p_gl_branch_rna

```

### Basal --> LP

```{r}
#| label: plot_dpt_blp

cells_use <- WhichCells(seurat, expression = !is.na(DPT_basal_lp) & is.finite(DPT_basal_lp) & annot_k4_rna_facs != "HS"  & annot_k4_rna_facs != "luminal")

# here luminal = LP by joint annot by selection
p_blp_facs <-RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_basal_lp",
  group.by = "facs_phenotype",
  cols = c("#ba8cba", "#91D1C2E5", "#E64B35E5"),
  sort = TRUE
)

p_blp_branch <-RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_basal_lp",
  group.by = "DPT_branch_label",
  cols = cols,
  sort = TRUE
)

p_blp_facs
#p_blp_branch


```



```{r}
#| label: plot_dpt_blp2

cells_use <- WhichCells(seurat, expression = !is.na(DPT_basal_lp_rna) & is.finite(DPT_basal_lp_rna) & annot_k4_rna_facs != "HS"  & annot_k4_rna_facs != "luminal")

# here luminal = LP by joint annot by selection
p_blp_k4_rna <- RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_basal_lp_rna",
  group.by = "annot_k4_rna_facs",
  cols = cols,
  sort = TRUE
) + coord_cartesian(xlim = c(0, 3.2))

p_blp_facs_rna <- RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_basal_lp_rna",
  group.by = "facs_phenotype",
  cols = c("#ba8cba", "#91D1C2E5", "#E64B35E5"),
  sort = TRUE
) + coord_cartesian(xlim = c(0, 3.2))


p_blp_facs_rna

```

```{r}
#| label: save_ridge
#| eval: false

ggsave(file.path(outputDir_plots, "ridge_blp_facs.pdf"),
       plot = p_blp_facs,
       device = "pdf",
       units = "mm",
       width = 180,
       height =100)

ggsave(file.path(outputDir_plots, "ridge_blp_facs_rna.pdf"),
       plot = p_blp_facs_rna,
       device = "pdf",
       units = "mm",
       width = 180,
       height =100)

```



# Transition characterisation
## Define transitioning cells 
```{r}
#| label: trans_define

pt <- seurat$DPT_basal_lp
lbl <- seurat$annot_k4_rna_facs
facs <- seurat$facs_phenotype

q_low  <- 0.75  # basal upper tail
q_high <- 0.10  # luminal lower tail

lb <- quantile(pt[lbl == "basal"], probs = q_low,  na.rm = TRUE)
ub <- quantile(pt[lbl == "LP"], probs = q_high, na.rm = TRUE)

c(lb = lb, ub = ub)

seurat$transition_annot <- dplyr::case_when(
  facs == "all_epith" & is.finite(pt) & pt >= lb & pt <= ub ~ "transitioning",
  facs == "all_epith" & pt < lb ~ "basal",
  facs == "all_epith" & pt > ub ~ "LP",
  lbl == "basal"  ~ "basal",
  lbl == "LP"  ~ "LP",
  lbl == "luminal" ~ "luminal",
  lbl == "HS" ~ "HS"
)

table(seurat$transition_annot)

```


```{r}
#| label: trans_ridge

cells_use <- WhichCells(seurat, expression = !is.na(DPT_basal_lp) & is.finite(DPT_basal_lp) & !is.na(transition_annot) & transition_annot != "luminal" & transition_annot != "HS")

p_ridge_trans <-RidgePlot(
  subset(seurat, cells = cells_use),
  features = "DPT_basal_lp",
  group.by = "transition_annot",
  cols = cols,
  sort = TRUE
)
p_ridge_trans

```

```{r}
#| label: save_ridge_trans
#| eval: false

ggsave(file.path(outputDir_plots, "ridge_blp_trans.pdf"),
       plot = p_ridge_trans,
       device = "pdf",
       units = "mm",
       width = 180,
       height =100)

```



```{r}
#| label: trans_plot

emb <- Seurat::Embeddings(seurat, reduction = "diffusion_spectral_k4")[, 1:3, drop = FALSE]
df <- data.frame(
  cell = rownames(emb),
  DC1  = emb[, 1],
  DC2  = emb[, 2],
  DC3  = emb[, 3]
)

df <- cbind(
  df,
  seurat@meta.data[df$cell, , drop = FALSE]
)
```

```{r}
#| label: trans_plot2

df_rest <- df[df$transition_annot != "transitioning", ]
df_trans <- df[df$transition_annot == "transitioning", ]

p_2 <- ggplot() +
        geom_point(
          data = df_rest,
          aes(DC1, DC2, color = transition_annot),
          size = 1.8,
          alpha = 1
        ) +
        geom_point(
          data = df_trans,
          aes(DC1, DC2, color = transition_annot),
          size = 1.8,
          alpha = 1
        ) +
        scale_color_manual(
          name = "transition_annot",
          values = cols,
          na.value = "grey80"
        ) +
        theme_classic() +
        labs(
          title = "diffusion_spectral_k4 by transition_annot",
          x = "DC1",
          y = "DC2"
        )

p_2
```

```{r}
#| label: trans_plot_save
#| eval: false

ggsave(file.path(outputDir_plots, "diff_map_trans.pdf"),
       plot = p_2,
       device = "pdf",
       units = "mm",
       width = 115,
       height =100)

```


# Motif calculation
```{r}
#| label: motif_calc
#| eval: false


library(TFBSTools)
library(BSgenome.Mmusculus.UCSC.mm10)
library(JASPAR2020)
library(chromVARmotifs)

data("mouse_pwms_v2")

pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection = "CORE",
              tax_group = 'vertebrates',
              species = "Mus musculus",  
              all_versions = FALSE)
)

pwm_jaspar_mouse <- toPWM(pfm)
pwm_all <- c(pwm_jaspar_mouse, mouse_pwms_v2)

seurat <- AddMotifs(
  object = seurat,
  assay = "bin_5k_h3k4me1",
  genome = BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwm_all
)

```


#  Reload step 6

```{r}
#| label: save_step6
#| eval: false

saveRDS(seurat, file.path(outputDir_objects, "seurat_step6.rds"))

```

```{r}
#| label: load_step6

seurat <- readRDS(file.path(outputDir_objects, "seurat_step6.rds"))

```


# Markers of the transition
## RNA
```{r}
#| label: markers1

DefaultAssay(seurat) <- "SCT_fl"
Idents(seurat) <- "transition_annot"

trans_markers <- FindMarkers(seurat, ident.1 = "transitioning")
trans_vs_lp_markers <- FindMarkers(seurat, ident.1 = "transitioning", ident.2 = "LP")
trans_vs_basal_markers <- FindMarkers(seurat, ident.1 = "transitioning", ident.2 = "basal")

```

```{r}
#| label: markers2

get_deg_sets <- function(df, padj_cutoff = 0.05, lfc_cutoff = 1) {
  df <- df %>% rownames_to_column("gene")

  list(
    up = df %>%
      filter(p_val_adj < padj_cutoff, avg_log2FC >  lfc_cutoff) %>%
      pull(gene) %>% unique(),

    down = df %>%
      filter(p_val_adj < padj_cutoff, avg_log2FC < -lfc_cutoff) %>%
      pull(gene) %>% unique()
  )
}

sets_lp      <- get_deg_sets(trans_vs_lp_markers)
sets_basal   <- get_deg_sets(trans_vs_basal_markers)
sets_all <- get_deg_sets(trans_markers)

venn_up <- list(
  "Trans vs Basal"   = sets_basal$up,
  "Trans vs LP"      = sets_lp$up,
  "Trans vs all" = sets_all$up
)

venn_down <- list(
  "Trans vs Basal"   = sets_basal$down,
  "Trans vs LP"      = sets_lp$down,
  "Trans vs all" = sets_all$down
)


p_up <- ggvenn(
  venn_up,
  show_percentage = FALSE,
  stroke_size = 0.8,
  fill_color = c("#E64B35E5", "#ba8cba", "grey"),
  set_name_size = 4
) +
  ggtitle("UP genes (padj < 0.05, log2fold > 1)") +
  theme_classic()

p_down <- ggvenn(
  venn_down,
  show_percentage = FALSE,
  stroke_size = 0.8,
  fill_color = c("#E64B35E5", "#ba8cba", "grey"),
  set_name_size = 4
) +
  ggtitle("DOWN genes (padj < 0.05, log2fold < -1)") +
  theme_classic()

p_up
p_down


up_all3   <- Reduce(intersect, venn_up)
down_all3 <- Reduce(intersect, venn_down)
up_only_basal <- setdiff(sets_basal$up, union(sets_lp$up, sets_all$up))
up_only_lp <- setdiff(sets_lp$up, union(sets_basal$up, sets_all$up))

```

```{r}
#| label: markers1_save
#| eval: false 

write.csv(trans_markers, file.path(outputDir_tables, "trans_markers_RNA.csv"))
write.csv(trans_vs_lp_markers, file.path(outputDir_tables, "trans_vs_lp_markers_RNA.csv"))
write.csv(trans_vs_basal_markers, file.path(outputDir_tables, "trans_vs_basal_markers_RNA.csv"))

```

```{r}
#| label: markers1_save2
#| eval: false 

ggsave(file.path(outputDir_plots, "venn_up.pdf"),
       plot = p_up,
       device = "pdf",
       units = "mm",
       width = 80,
       height = 80)

ggsave(file.path(outputDir_plots, "venn_down.pdf"),
       plot = p_down,
       device = "pdf",
       units = "mm",
       width = 80,
       height = 80)
```


## H3K4me1 GA
```{r}
#| label: markers_k4

DefaultAssay(seurat) <- "ga_h3k4me1"
Idents(seurat) <- "transition_annot"

trans_markers_k4 <- FindMarkers(seurat, ident.1 = "transitioning")
trans_vs_lp_markers_k4 <- FindMarkers(seurat, ident.1 = "transitioning", ident.2 = "LP")
trans_vs_basal_markers_k4 <- FindMarkers(seurat, ident.1 = "transitioning", ident.2 = "basal")

```

```{r}
#| label: markers_k4_2

sets_lp      <- get_deg_sets(trans_vs_lp_markers_k4)
sets_basal   <- get_deg_sets(trans_vs_basal_markers_k4)
sets_all <- get_deg_sets(trans_markers_k4)

venn_up <- list(
  "Trans vs Basal"   = sets_basal$up,
  "Trans vs LP"      = sets_lp$up,
  "Trans vs all" = sets_all$up
)

venn_down <- list(
  "Trans vs Basal"   = sets_basal$down,
  "Trans vs LP"      = sets_lp$down,
  "Trans vs all" = sets_all$down
)


p_up_k4 <- ggvenn(
  venn_up,
  show_percentage = FALSE,
  stroke_size = 0.8,
  fill_color = c("#E64B35E5", "#ba8cba", "grey"),
  set_name_size = 4
) +
  ggtitle("UP GA K4 (padj < 0.05, log2fold > 1)") +
  theme_classic()

p_down_k4 <- ggvenn(
  venn_down,
  show_percentage = FALSE,
  stroke_size = 0.8,
  fill_color = c("#E64B35E5", "#ba8cba", "grey"),
  set_name_size = 4
) +
  ggtitle("DOWN GA K4 (padj < 0.05, log2fold < -1)") +
  theme_classic()

p_up_k4
p_down_k4

```



```{r}
#| label: markers2_2_save
#| eval: false

write.csv(trans_markers_k4, file.path(outputDir_tables, "trans_markers_ga_k4.csv"))
write.csv(trans_vs_lp_markers_k4, file.path(outputDir_tables, "trans_vs_lp_markers_ga_k4.csv"))
write.csv(trans_vs_basal_markers_k4, file.path(outputDir_tables, "trans_vs_basal_markers_ga_k4.csv"))

```

## H3K4me1 bins
```{r}
#| label: markers3
#| eval: false

DefaultAssay(seurat) <- "bin_5k_h3k4me1"
Idents(seurat) <- "transition_annot"

trans_markers_k4_bin <- FindMarkers(seurat, ident.1 = "transitioning")
trans_vs_lp_markers_k4_bin <- FindMarkers(seurat, ident.1 = "transitioning", ident.2 = "LP")
trans_vs_basal_markers_k4_bin <- FindMarkers(seurat, ident.1 = "transitioning", ident.2 = "basal")

```

```{r}
#| label: markers3_2
#| eval: false

write.csv(trans_markers_k4_bin, file.path(outputDir_tables, "trans_markers_bin_k4.csv"))
write.csv(trans_vs_lp_markers_k4_bin, file.path(outputDir_tables, "trans_vs_lp_markers_bin_k4.csv"))
write.csv(trans_vs_basal_markers_k4_bin, file.path(outputDir_tables, "trans_vs_basal_markers_bin_k4.csv"))

```

# Trans vs all 
## GSEA RNA 
```{r}
#| label: gsea_rna_1

rnk <- trans_markers %>%
  rownames_to_column("gene") %>%
  distinct(gene, .keep_all = TRUE)


geneList <- rnk$avg_log2FC
names(geneList) <- rnk$gene
geneList <- sort(geneList, decreasing = TRUE)

```

```{r}
#| label: gsea_rna_2

msig_h <- msigdbr(species = "Mus musculus", collection = "H")
term2gene_sym <- msig_h %>%
  dplyr::select(gs_name, gene_symbol) %>%
  distinct()

gsea_h <- GSEA(
  geneList      = geneList,
  TERM2GENE     = term2gene_sym,
  minGSSize     = 10,
  maxGSSize     = 500,
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  verbose       = FALSE
)

write.csv(gsea_h, file.path(outputDir_tables, "gsea_hallmark.csv"))
```


```{r}
#| label: gsea_rna_3

plot_gsea_nes <- function(gsea_res,
                         title = NULL,
                         n_each = 5,
                         padj_cutoff = 0.05,
                         wrap_width = 45) {
  df <- as.data.frame(gsea_res)

  df_sig <- df %>%
    filter(!is.na(p.adjust), p.adjust <= padj_cutoff, !is.na(NES)) %>%
    mutate(Direction = if_else(NES >= 0, "pos", "neg"))

  if (nrow(df_sig) == 0) {
    return(ggplot() + theme_void() + ggtitle(paste0(title, " (no significant terms)")))
  }

  # top N positive and top N negative (by smallest p.adjust)
  top_pos <- df_sig %>% filter(Direction == "pos") %>% arrange(p.adjust) %>% slice_head(n = n_each)
  top_neg <- df_sig %>% filter(Direction == "neg") %>% arrange(p.adjust) %>% slice_head(n = n_each)

  df_top <- bind_rows(top_pos, top_neg) %>%
    mutate(
      Description = str_wrap(Description, width = wrap_width),
      # order terms by NES so the y-axis is meaningful
      Description = reorder(Description, NES)
    )

  # choose the "count" column: in GSEA results, setSize is the gene set size
  count_col <- if ("Count" %in% names(df_top)) "Count" else if ("setSize" %in% names(df_top)) "setSize" else NULL
  if (is.null(count_col)) {
    stop("No Count or setSize column found in the GSEA result table.")
  }

  ggplot(df_top, aes(x = NES, y = Description)) +
    geom_point(
        aes(size = .data[[count_col]], fill = p.adjust),
        shape = 21,
        color = "black",
        stroke = 0.6
      ) +
    ggtitle(title) +
    xlab("NES") +
    ylab(NULL) +
    scale_size_continuous(name = count_col) +
    scale_fill_gradient(
       low = mypal[1],
       high = "grey",
       name = "p.adjust"
     ) +
    theme_bw() +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4)
}

```

```{r}
#| label: gsea_rna_4

p_h  <- plot_gsea_nes(gsea_h,  title = "MSigDB hallmark GSEA", n_each = 5)
p_h

```

```{r}
#| label: gsea_rna_save
#| eval: false

ggsave(file.path(outputDir_plots, "gsea_h.pdf"),
       plot = p_h,
       device = "pdf",
       units = "mm",
       width = 150,
       height = 80)
```

## GSEA K4
```{r}
#| label: gsea_k4

rnk <- trans_markers_k4 %>%
  rownames_to_column("gene") %>%
  distinct(gene, .keep_all = TRUE)


geneList <- rnk$avg_log2FC
names(geneList) <- rnk$gene
geneList <- sort(geneList, decreasing = TRUE)

```


```{r}
#| label: gsea_k4_2

msig_h <- msigdbr(species = "Mus musculus", collection = "H")
term2gene_sym <- msig_h %>%
  dplyr::select(gs_name, gene_symbol) %>%
  distinct()

gsea_h_k4 <- GSEA(
  geneList      = geneList,
  TERM2GENE     = term2gene_sym,
  minGSSize     = 10,
  maxGSSize     = 500,
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  verbose       = FALSE
)

write.csv(gsea_h_k4, file.path(outputDir_tables, "gsea_hallmark_k4.csv"))
```


```{r}
#| label: gsea_k4_3

p_h_k4  <- plot_gsea_nes(gsea_h_k4,  title = "MSigDB hallmark GSEA K4 GA", n_each = 5)
p_h_k4

```

```{r}
#| label: gsea_k4_4
#| eval: false


ggsave(file.path(outputDir_plots, "gsea_h_k4.pdf"),
       plot = p_h_k4,
       device = "pdf",
       units = "mm",
       width = 150,
       height = 50)
```

## Volc
### RNA
```{r}
#| label: volc_rna

df_volc <- trans_markers
df_volc$gene <- rownames(df_volc)

df_volc$neglog10_padj <- -log10(df_volc$p_val_adj)


top_genes <- subset(
  df_volc,
  p_val_adj < 1e-6 & abs(avg_log2FC) > 2
)

p_volcano <- ggplot(df_volc, aes(x = avg_log2FC, y = neglog10_padj)) +
  geom_point(color = "grey60", size = 1.6) +
  geom_point(
    data = subset(df_volc, p_val_adj < 0.05 & abs(avg_log2FC) > 1),
    color = mypal[5],
    size = 1.8
  ) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  theme_classic() +
  labs(
    x = "avg log2FC",
    y = "-log10 adjusted p-value",
    title = "Transitioning vs all"
  )

p_volcano

```

```{r}
#| label: volc_rna_save
#| eval: false

ggsave(file.path(outputDir_plots, "volc_trans_vs_all_rna.pdf"),
       plot = p_volcano,
       device = "pdf",
       units = "mm",
       width = 80,
       height = 100)

```


### K4
```{r}
#| label: volc_k4

df_volc_k4 <- trans_markers_k4

# remove absurd points with -200 fold and padj = 1
df_volc_k4 <- trans_markers_k4 %>%
  dplyr::filter(p_val_adj < 0.6)
df_volc_k4$gene <- rownames(df_volc_k4)

df_volc_k4$neglog10_padj <- -log10(df_volc_k4$p_val_adj)


top_genes <- subset(
  df_volc_k4,
  p_val_adj < 0.05 & abs(avg_log2FC) > 1
)

p_volcano_k4 <- ggplot(df_volc_k4, aes(x = avg_log2FC, y = neglog10_padj)) +
  geom_point(color = "grey60", size = 1.6) +
  geom_point(
    data = subset(df_volc_k4, p_val_adj < 0.05 & abs(avg_log2FC) > 1),
    color = mypal[2],
    size = 1.8
  ) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  theme_classic() +
  labs(
    x = "avg log2FC",
    y = "-log10 adjusted p-value",
    title = "Transitioning vs all, K4 GA"
  )

p_volcano_k4

```


```{r}
#| label: volc_k4_save
#| eval: false

ggsave(file.path(outputDir_plots, "volc_trans_vs_all_k4.pdf"),
       plot = p_volcano_k4,
       device = "pdf",
       units = "mm",
       width = 80,
       height = 100)

```


# Trans vs basal
## GSEA RNA 
```{r}
#| label: gsea_rna_vsbasal1

rnk <- trans_vs_basal_markers %>%
  rownames_to_column("gene") %>%
  distinct(gene, .keep_all = TRUE)

geneList <- rnk$avg_log2FC
names(geneList) <- rnk$gene
geneList <- sort(geneList, decreasing = TRUE)

```


```{r}
#| label: gsea_rna_vsbasal2

msig_h <- msigdbr(species = "Mus musculus", collection = "H")
term2gene_sym <- msig_h %>%
  dplyr::select(gs_name, gene_symbol) %>%
  distinct()

gsea_h_vs_basal <- GSEA(
  geneList      = geneList,
  TERM2GENE     = term2gene_sym,
  minGSSize     = 10,
  maxGSSize     = 500,
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  verbose       = FALSE
)

write.csv(gsea_h_vs_basal, file.path(outputDir_tables, "gsea_hallmark_vs_basal.csv"))
```

```{r}
#| label: gsea_rna_vsbasal3

p_h_vs_basal  <- plot_gsea_nes(gsea_h_vs_basal,  title = "MSigDB hallmark GSEA", n_each = 5)
p_h_vs_basal

```

```{r}
#| label: gsea_rna_vsbasal_save
#| eval: false

ggsave(file.path(outputDir_plots, "gsea_h_vs_basal.pdf"),
       plot = p_h_vs_basal,
       device = "pdf",
       units = "mm",
       width = 140,
       height = 80)

```

## GSEA K4
```{r}
#| label: gsea_k4_vsbasal

rnk <- trans_vs_basal_markers_k4 %>%
  rownames_to_column("gene") %>%
  distinct(gene, .keep_all = TRUE)


geneList <- rnk$avg_log2FC
names(geneList) <- rnk$gene
geneList <- sort(geneList, decreasing = TRUE)

```


```{r}
#| label: gsea_k4_vsbasal2

msig_h <- msigdbr(species = "Mus musculus", collection = "H")
term2gene_sym <- msig_h %>%
  dplyr::select(gs_name, gene_symbol) %>%
  distinct()

gsea_h_k4_vs_basal <- GSEA(
  geneList      = geneList,
  TERM2GENE     = term2gene_sym,
  minGSSize     = 10,
  maxGSSize     = 500,
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  verbose       = FALSE
)

write.csv(gsea_h_k4_vs_basal, file.path(outputDir_tables, "gsea_hallmark_k4_vs_basal.csv"))
```



```{r}
#| label: gsea_k4_vsbasal3

p_h_k4_vs_basal  <- plot_gsea_nes(gsea_h_k4_vs_basal,  title = "MSigDB hallmark GSEA K4 GA", n_each = 5)
p_h_k4_vs_basal

```

```{r}
#| label: gsea_k4_vsbasal_save
#| eval: false

ggsave(file.path(outputDir_plots, "gsea_h_k4_vs_basal.pdf"),
       plot = p_h_k4_vs_basal,
       device = "pdf",
       units = "mm",
       width = 150,
       height = 50)

```

## Volc
### RNA
```{r}
#| label: volc_rna_vsbasal

df_volc <- trans_vs_basal_markers
df_volc$gene <- rownames(df_volc)

df_volc$neglog10_padj <- -log10(df_volc$p_val_adj)


top_genes <- df_volc |>
  dplyr::filter(p_val_adj < 0.05, abs(avg_log2FC) > 1) |>
  dplyr::mutate(direction = ifelse(avg_log2FC > 0, "UP", "DOWN")) |>
  dplyr::group_by(direction) |>
  dplyr::arrange(p_val_adj, .by_group = TRUE) |>
  dplyr::slice_head(n = 16) |>
  dplyr::ungroup()

p_volcano_trans_vs_basal <- ggplot(df_volc, aes(x = avg_log2FC, y = neglog10_padj)) +
  geom_point(color = "grey60", size = 1.6) +
  geom_point(
    data = subset(df_volc, p_val_adj < 0.05 & abs(avg_log2FC) > 1),
    color = mypal[5],
    size = 1.8
  ) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  theme_classic() +
  labs(
    x = "avg log2FC",
    y = "-log10 adjusted p-value",
    title = "Transitioning vs basal"
  )

p_volcano_trans_vs_basal

```

```{r}
#| label: volc_rna_vsbasal_save
#| eval: false

ggsave(file.path(outputDir_plots, "volc_trans_vs_basal_rna.pdf"),
       plot = p_volcano_trans_vs_basal,
       device = "pdf",
       units = "mm",
       width = 90,
       height = 100)

```


### K4

```{r}
#| label: volc_k4_vsbasal
# remove absurd points with -200 fold and padj = 1
df_volc_k4 <- trans_vs_basal_markers_k4 %>%
  dplyr::filter(p_val_adj < 0.2)
df_volc_k4$gene <- rownames(df_volc_k4)

df_volc_k4$neglog10_padj <- -log10(df_volc_k4$p_val_adj)


top_genes <- subset(
  df_volc_k4,
  p_val_adj < 0.05 & abs(avg_log2FC) > 1
)

p_volcano_k4_trans_vs_basal <- ggplot(df_volc_k4, aes(x = avg_log2FC, y = neglog10_padj)) +
  geom_point(color = "grey60", size = 1.6) +
  geom_point(
    data = subset(df_volc_k4, p_val_adj < 0.05 & abs(avg_log2FC) > 1),
    color = mypal[2],
    size = 1.8
  ) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  theme_classic() +
  labs(
    x = "avg log2FC",
    y = "-log10 adjusted p-value",
    title = "Transitioning vs all, K4 GA"
  )

p_volcano_k4_trans_vs_basal

```


```{r}
#| label: volc_k4_vsbasal_save
#| eval: false

ggsave(file.path(outputDir_plots, "volc_trans_vs_basal_k4.pdf"),
       plot = p_volcano_k4_trans_vs_basal,
       device = "pdf",
       units = "mm",
       width = 120,
       height = 100)

```

# Trans vs LP
## GSEA RNA 
```{r}
#| label: gsea_rna_vslp1

rnk <- trans_vs_lp_markers %>%
  rownames_to_column("gene") %>%
  distinct(gene, .keep_all = TRUE)

geneList <- rnk$avg_log2FC
names(geneList) <- rnk$gene
geneList <- sort(geneList, decreasing = TRUE)

```


```{r}
#| label: gsea_rna_vslp2

msig_h <- msigdbr(species = "Mus musculus", collection = "H")
term2gene_sym <- msig_h %>%
  dplyr::select(gs_name, gene_symbol) %>%
  distinct()

gsea_h_vs_lp <- GSEA(
  geneList      = geneList,
  TERM2GENE     = term2gene_sym,
  minGSSize     = 10,
  maxGSSize     = 500,
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  verbose       = FALSE
)

write.csv(gsea_h_vs_lp, file.path(outputDir_tables, "gsea_hallmark_vs_lp.csv"))
```



```{r}
#| label: gsea_rna_vslp3

p_h_vs_lp <- plot_gsea_nes(gsea_h_vs_lp,  title = "MSigDB hallmark GSEA", n_each = 5)

p_h_vs_lp
```

```{r}
#| label: gsea_rna_vslp3_save

ggsave(file.path(outputDir_plots, "gsea_h_vs_lp.pdf"),
       plot = p_h_vs_lp,
       device = "pdf",
       units = "mm",
       width = 138,
       height = 80)

```

## GSEA K4
```{r}
#| label: gsea_k4_vslp1

rnk <- trans_vs_lp_markers_k4 %>%
  rownames_to_column("gene") %>%
  distinct(gene, .keep_all = TRUE)


geneList <- rnk$avg_log2FC
names(geneList) <- rnk$gene
geneList <- sort(geneList, decreasing = TRUE)

```


```{r}
#| label: gsea_k4_vslp2

msig_h <- msigdbr(species = "Mus musculus", collection = "H")
term2gene_sym <- msig_h %>%
  dplyr::select(gs_name, gene_symbol) %>%
  distinct()

gsea_h_k4_vs_lp <- GSEA(
  geneList      = geneList,
  TERM2GENE     = term2gene_sym,
  minGSSize     = 10,
  maxGSSize     = 500,
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  verbose       = FALSE
)

write.csv(gsea_h_k4_vs_lp, file.path(outputDir_tables, "gsea_hallmark_k4_vs_lp.csv"))
```



```{r}
#| label: gsea_k4_vslp3

p_h_k4_vs_lp  <- plot_gsea_nes(gsea_h_k4_vs_lp,  title = "MSigDB hallmark GSEA K4 GA", n_each = 5)
p_h_k4_vs_lp

```

```{r}
#| label: gsea_k4_vslp3_save
#| eval: false

ggsave(file.path(outputDir_plots, "gsea_h_k4_vs_lp.pdf"),
       plot = p_h_k4_vs_lp,
       device = "pdf",
       units = "mm",
       width = 150,
       height = 50)

```

## Volc
### RNA
```{r}
#| label: volc_rna_vslp

df_volc <- trans_vs_lp_markers
df_volc$gene <- rownames(df_volc)

df_volc$neglog10_padj <- -log10(df_volc$p_val_adj)

top_genes <- df_volc |>
  dplyr::filter(p_val_adj < 0.05, abs(avg_log2FC) > 1) |>
  dplyr::mutate(direction = ifelse(avg_log2FC > 0, "UP", "DOWN")) |>
  dplyr::group_by(direction) |>
  dplyr::arrange(p_val_adj, .by_group = TRUE) |>
  dplyr::slice_head(n = 17) |>
  dplyr::ungroup()

p_volcano_trans_vs_lp <- ggplot(df_volc, aes(x = avg_log2FC, y = neglog10_padj)) +
  geom_point(color = "grey", size = 1.6) +
  geom_point(
    data = subset(df_volc, p_val_adj < 0.05 & abs(avg_log2FC) > 1),
    color = mypal[5],
    size = 1.8
  ) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey") +
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  theme_classic() +
  labs(
    x = "avg log2FC",
    y = "-log10 adjusted p-value",
    title = "Transitioning vs LP"
  )

p_volcano_trans_vs_lp

```

```{r}
#| label: volc_rna_vslp_save
#| eval: false

ggsave(file.path(outputDir_plots, "volc_trans_vs_lp_rna.pdf"),
       plot = p_volcano_trans_vs_lp,
       device = "pdf",
       units = "mm",
       width = 90,
       height = 100)

```


### K4

```{r}
#| label: volc_k4_vslp

# remove absurd points with -200 fold and padj = 1
df_volc_k4 <- trans_vs_lp_markers_k4 %>%
  dplyr::filter(p_val_adj < 1)
df_volc_k4$gene <- rownames(df_volc_k4)

df_volc_k4$neglog10_padj <- -log10(df_volc_k4$p_val_adj)


top_genes <- subset(
  df_volc_k4,
  p_val_adj < 0.05 & abs(avg_log2FC) > 1
)

p_volcano_k4_trans_vs_lp <- ggplot(df_volc_k4, aes(x = avg_log2FC, y = neglog10_padj)) +
  geom_point(color = "grey60", size = 1.6) +
  geom_point(
    data = subset(df_volc_k4, p_val_adj < 0.05 & abs(avg_log2FC) > 1),
    color = mypal[2],
    size = 1.8
  ) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  theme_classic() +
  labs(
    x = "avg log2FC",
    y = "-log10 adjusted p-value",
    title = "Transitioning vs LP, K4 GA"
  )

p_volcano_k4_trans_vs_lp

```

```{r}
#| label: volc_k4_vslp_save
#| eval: false

ggsave(file.path(outputDir_plots, "volc_trans_vs_lp_k4.pdf"),
       plot = p_volcano_k4_trans_vs_lp,
       device = "pdf",
       units = "mm",
       width = 120,
       height = 100)

```


