---
title: "Functions"
author: "Anna Schwager"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Weighted histograms
This functions plots a weighted histogram from a Seurat object. 
```{r weighted_hist}
#' Plot a weighted histogram of log10-transformed reads per cell
#'
#' This function generates a weighted histogram of log10-transformed reads per cell,
#' with an option to switch between RNA assay (`nCount_RNA`) and chromatin assays (`reads_count`).
#'
#' @param seurat A Seurat object containing single-cell data.
#' @param assay A character string indicating the assay to use. Default is the current default assay
#'   in the Seurat object. If "RNA", the function will use the column `nCount_RNA` from the metadata,
#'   otherwise it will use the `reads_count` column.
#' 
#' @return A `ggplot2` object showing a weighted histogram.
#' 
#' @examples
#' # Assuming 'seurat_k27' is a Seurat object with 'reads_count' (calculated using the
#' CountFragments function of Signac) and 'nCount_RNA' in meta.data
#' plot_weighted_hist(seurat_k27)
#' plot_weighted_hist(seurat_k27, assay = "RNA")
#'
plot_weighted_hist = function(seurat, assay = DefaultAssay(seurat)) {
  # Determine which column to use based on the assay
  if (assay == "RNA") {
    count_column <- "nCount_RNA"
    fill="lightgoldenrod"
  } else if (assay == "RNA_features") {
    count_column <- "nFeature_RNA"
    fill="gold"
  } else if (assay == "RNA_umi"){
    count_column <- "nCount_RNA_umi"
    fill="lightgoldenrod"
  } else if (assay == "RNA_fl"){
    count_column <- "nCount_RNA_fl"
    fill="lightgoldenrod"
  } else if (assay == "RNA_features_umi"){
    count_column <- "nFeature_RNA_umi"
    fill="gold"
  } else if (assay == "RNA_features_fl"){
    count_column <- "nFeature_RNA_fl"
    fill="gold"
  } else {
    count_column <- "frequency_count"
    fill="lightskyblue2"
  }
  
  # Extract the appropriate counts from the meta data
  counts <- seurat@meta.data[[count_column]]
  log_counts <- log10(counts)
  
  # Create the plot
  ggplot(data.frame(log_counts), aes(x=log_counts)) +
    geom_histogram(aes(weight = log_counts), color="black", fill=fill) +
    ylab("N barcodes") +
    xlab("log10(reads/cell)") + 
    theme_minimal() + 
    theme(text = element_text(size=20))
}

```

# Cell type scores 
This function calculates the cell type scores = average expression or gene activity of a given set of reference genes for each cell
```{r score_function, message=F, warning=F}

AddCellTypeScore <- function(seurat_obj,
                             marker_table,
                             cell_type,
                             assay = "SCT_fl",
                             slot = "data",
                             score_name = "cell_type_score") {
  # Get marker genes for specified cell type
  marker_genes <- marker_table %>%
    filter(Type == cell_type) %>%
    pull(Gene) %>%
    unique()
  
  # Extract expression matrix
  expr_matrix <- GetAssayData(seurat_obj, assay = assay, layer = slot)
  
  marker_genes_present <- intersect(marker_genes, rownames(expr_matrix))
  
  if (length(marker_genes_present) == 0) {
    warning(paste0("No marker genes found in assay for cell type: ", cell_type))
    seurat_obj[[score_name]] <- NA
    return(seurat_obj)
  }
  
  score_vector <- Matrix::colMeans(expr_matrix[marker_genes_present, , drop = FALSE])
  
  seurat_obj[[score_name]] <- score_vector
  
  return(seurat_obj)
}

```


# Mouse to human conversion
These functions convert human genes to mouse orthologs and vica versa.
Adapted from the Biostars post of Nitin Narwade 

```{r mouse_human}

convert_mouse_to_human <- function(gene_list) { 
     output = c()
     mouse_human_genes = read.csv("https://www.informatics.jax.org/downloads/reports/HOM_MouseHumanSequence.rpt",sep="\t")

     for(gene in gene_list) {
          class_key = (mouse_human_genes %>% filter(Symbol == gene & Common.Organism.Name == "mouse, laboratory"))[['DB.Class.Key']]
          if( !identical(class_key, integer(0)) ) {
               human_genes = (mouse_human_genes %>% filter(DB.Class.Key == class_key & Common.Organism.Name=="human"))[,"Symbol"]
               for(human_gene in human_genes) {
                    output = rbind(c(gene, human_gene), output)
               }
          }
     }
     return (output)
}

convert_human_to_mouse <- function(gene_list) {
    output = c()
    mouse_human_genes = read.csv("https://www.informatics.jax.org/downloads/reports/HOM_MouseHumanSequence.rpt",sep="\t")

    for(gene in gene_list) {
          class_key = (mouse_human_genes %>% filter(Symbol == gene & Common.Organism.Name == "human"))[['DB.Class.Key']]
          if( !identical(class_key, integer(0)) ) {
            human_genes = (mouse_human_genes %>% filter(DB.Class.Key == class_key & Common.Organism.Name=="mouse, laboratory"))[,"Symbol"]
            for(human_gene in human_genes) {
                output = rbind(c(gene, human_gene), output)
            }
          }
     }
     return (output)
}
```


# Create bigwigs per cluster 
This function (and its helpers) exports bigwigs per cluster.
Taken from (https://github.com/lldelisle/signac/blob/export_group_bw_rebased/R/utilities.R)

```{r export_bw}
#' Export bigwig files for groups of cells
#'
#' @param object A Seurat object
#' @param assay Name of assay to use
#' @param group.by The metadata variable used to group the cells
#' @param idents Identities to include (defined by group.by parameter)
#' @param normMethod Normalization method for the bigwig files. Default 'RC'.
#' 'RC' will divide the number of fragments in a tile by the total number of
#' fragments in the group. A scaling factor of 10^4 will be applied. 
#' 'ncells' will divide the number of fragments in a tile by the number of cell
#' in the group. 'none' will apply no normalization method. 
#' A vector of values for each cell can also be passed as a metadata column
#' name. A scaling factor of 10^4 will be applied
#' @param tileSize The size of the tiles in the bigwig file
#' @param minCells The minimum of cells in a group to be exported
#' @param cutoff The maximum number of fragments in a given genomic tile
#' @param chromosome A chromosomes vector to be exported
#' @param outdir Directory to write output files (splitted bed and bigwigs)
#' @param verbose Display messages
#'
#' @importFrom GenomicRanges GRanges slidingWindows seqlevels seqlengths coverage
#' @importFrom IRanges IRanges
#' @importFrom future nbrOfWorkers
#' @importFrom future.apply future_lapply
#' @importFrom pbapply pblapply
#' @importFrom rtracklayer import export.bw
#' @importFrom BiocGenerics which
#' @importFrom S4Vectors match
#' @importFrom Matrix sparseMatrix rowSums
#' @importFrom SeuratObject Idents
#'
#' @export
#'
#' @examples
#' \dontrun{
#' ExportGroupBW(object, assay = "peaks")
#' }
ExportGroupBW  <- function(
    object,
    assay = NULL,
    group.by = NULL,
    idents = NULL,
    normMethod = "RC",
    tileSize = 100,
    minCells = 5,
    cutoff = NULL,
    chromosome = NULL,
    outdir = getwd(),
    verbose = TRUE
) {
  # Check output dir
  if (!dir.exists(outdir)){
    dir.create(outdir, recursive = TRUE)
  }
  # Fragments present?
  if (length(Fragments(object)) == 0) {
    stop("This object does not have Fragments, cannot generate bigwig.")
  }
  # Restrict chromosomes if provided
  if(!is.null(x = chromosome)){
    seqlevels(object) <- chromosome
  }
  chromLengths <- seqlengths(object)
  if (is.null(chromLengths)) {
    stop("Object has no seqlength, bigwig coverages cannot be evaluated.")
  }
  availableChr <- names(chromLengths)
  chromSizes <- GRanges(
    seqnames = availableChr,
    ranges = IRanges(
      start = rep(1, length(x = availableChr)),
      end = as.numeric(x = chromLengths)
    )
  )
  assay <- SetIfNull(x = assay, y = DefaultAssay(object = object))
  obj.groups <- GetGroups(
    object = object,
    group.by = group.by,
    idents = idents
  )
  GroupsNames <- names(x = table(obj.groups)[table(obj.groups) > minCells])

  # Warn/overwrite pre-existing .bed
  lapply(X = GroupsNames, FUN = function(x) {
    fn <- paste0(outdir, .Platform$file.sep, x, ".bed")
    if (file.exists(fn)) {
      message(sprintf("The group \"%s\" is already present in the destination folder and will be overwritten!", x))
      file.remove(fn)
    }
  })

  # Split fragments per group
  SplitFragments(
    object = object,
    assay = assay,
    group.by = group.by,
    idents = idents,
    outdir = outdir,
    file.suffix = "",
    append = TRUE,
    buffer_length = 256L,
    verbose = verbose
  )

  # Column to normalize by
  if(!is.null(x = normMethod)) {
    if (tolower(x = normMethod) %in% c('rc', 'ncells', 'none')){
      normBy <- normMethod
    } else{
      normBy <- object[[normMethod, drop = FALSE]]
    }
  } else {
    normBy <- NULL
  }

  if (verbose) message("Creating tiles")
  tiles <- unlist(
    x = slidingWindows(x = chromSizes, width = tileSize, step = tileSize)
  )

  if (verbose) message("Creating bigwig files at ", outdir)

  # Choose apply backend
  if (future::nbrOfWorkers() > 1) {
    mylapply <- future.apply::future_lapply
  } else {
    if (isTRUE(verbose) && requireNamespace("pbapply", quietly = TRUE)) {
      mylapply <- pbapply::pblapply
    } else {
      if (isTRUE(verbose) && !requireNamespace("pbapply", quietly = TRUE)) {
        message("Package 'pbapply' not available; using base lapply (no progress bar).")
      }
      mylapply <- lapply
    }
  }

  covFiles <- mylapply(
    GroupsNames,
    FUN = CreateBWGroup,
    availableChr,
    chromLengths,
    tiles,
    normBy,
    tileSize,
    normMethod,
    cutoff,
    outdir
  )
  return(covFiles)
}

# Helper: create a group bigwig
CreateBWGroup <- function(
    groupNamei,
    availableChr,
    chromLengths,
    tiles,
    normBy,
    tileSize,
    normMethod,
    cutoff,
    outdir
) {
  normMethod <- tolower(x = normMethod)

  # Read group fragments (.bed) produced by SplitFragments()
  fragi <- rtracklayer::import(
    paste0(outdir, .Platform$file.sep, groupNamei, ".bed"),
    format = "bed"
  )

  cellGroupi <- unique(x = fragi$name)

  # Output file
  covFile <- file.path(
    outdir,
    paste0(groupNamei, "-TileSize-", tileSize, "-normMethod-", normMethod, ".bw")
  )

  covList <- lapply(X = seq_along(availableChr), FUN = function(k) {
    fragik <- fragi[seqnames(fragi) == availableChr[k], ]
    tilesk <- tiles[BiocGenerics::which(S4Vectors::match(seqnames(tiles), availableChr[k], nomatch = 0) > 0)]

    if (length(x = fragik) == 0) {
      tilesk$reads <- 0
    } else {
      # Number of tiles for this chromosome
      nTiles <- as.numeric(chromLengths[availableChr[k]]) / tileSize
      if (nTiles %% 1 != 0) nTiles <- trunc(nTiles) + 1
      nTiles <- as.integer(nTiles)

      # Map fragment names to columns; drop non-matching
      matchID <- S4Vectors::match(mcols(fragik)$name, cellGroupi)

      # Compute start/end tile indices (1-based), clamp to [1, nTiles]
      start_idx <- trunc((start(fragik) - 1L) / tileSize) + 1L
      end_idx   <- trunc((end(fragik)   - 1L) / tileSize) + 1L
      start_idx <- pmin(pmax(start_idx, 1L), nTiles)
      end_idx   <- pmin(pmax(end_idx,   1L), nTiles)

      # Keep only valid columns
      ok <- !is.na(matchID)

      if (!any(ok)) {
        tilesk$reads <- 0
      } else {
        i <- c(start_idx[ok], end_idx[ok])
        j <- rep.int(as.integer(matchID[ok]), 2L)

        mat <- Matrix::sparseMatrix(
          i = as.integer(i),
          j = as.integer(j),
          x = rep(1L, length(i)),
          dims = c(nTiles, length(cellGroupi))
        )

        # Cap counts per tile if requested
        if (!is.null(x = cutoff)) {
          mat@x[mat@x > cutoff] <- cutoff
        }

        # Sum cells → per-tile counts
        mat <- Matrix::rowSums(mat)
        tilesk$reads <- mat

        # Normalization
        if (!is.null(x = normMethod)) {
          if (normMethod == "rc") {
            tilesk$reads <- tilesk$reads * 10^4 / length(fragi$name)
          } else if (normMethod == "ncells") {
            tilesk$reads <- tilesk$reads / length(cellGroupi)
          } else if (normMethod == "none") {
            # no-op
          } else {
            # vector per cell passed via metadata column
            if (!is.null(x = normBy)) {
              tilesk$reads <- tilesk$reads * 10^4 / sum(normBy[cellGroupi, 1])
            }
          }
        }
      }
    }

    # Convert to coverage Rle for this chromosome
    tilesk <- coverage(tilesk, weight = tilesk$reads)[[availableChr[k]]]
    tilesk
  })

  names(covList) <- availableChr
  covList <- as(object = covList, Class = "RleList")
  rtracklayer::export.bw(object = covList, con = covFile)
  return(covFile)
}

# Set default if null
SetIfNull <- function(x, y) {
  if (is.null(x = x)) y else x
}

# Get vector of cell names and associated identity
GetGroups <- function(
    object,
    group.by,
    idents
) {
  if (is.null(x = group.by)) {
    obj.groups <- SeuratObject::Idents(object = object)
  } else {
    obj.md <- object[[group.by]]
    obj.groups <- obj.md[, 1]
    names(obj.groups) <- rownames(x = obj.md)
  }
  if (!is.null(idents)) {
    obj.groups <- obj.groups[obj.groups %in% idents]
  }
  return(obj.groups)
}

```


# Correlation-based filtering adapted from ChromScape

Modified ChromScape functions with Cosine distance instead of Pearson + choosing the reductions to use 
```{r chromscape1}
correlation_and_hierarchical_clust_scExp_cosine <- function(
    scExp, hc_linkage = "ward.D", reduction = "PCA"){
    
    stopifnot(is(scExp, "SingleCellExperiment"), is.character(hc_linkage))
    
    if (is.null(SingleCellExperiment::reducedDim(scExp, reduction)))
        stop(
            "ChromSCape::correlation_and_hierarchical_clust_scExp -
                Run PCA on the object before correlation")
    
    pca <- SingleCellExperiment::reducedDim(scExp, reduction)
    pca_t <- Matrix::t(pca)   # cells × PCs
    
    # --- NEW: cosine similarity instead of Pearson correlation ---
    cos_mat <- coop::cosine(pca_t)
    
    # convert to cosine distance
    dist_mat <- stats::as.dist(1 - cos_mat)
    
    # hierarchical clustering
    hc_cos <- stats::hclust(dist_mat, method = hc_linkage)
    hc_cos$labels <- rep("", ncol(scExp))
    
    # store results
    scExp@metadata$hc_cor <- hc_cos
    SingleCellExperiment::reducedDim(scExp, "Cor") <- as(cos_mat, "dspMatrix")
    
    return(scExp)
}

```


```{r chromscape2}
filter_correlated_cell_scExp <- function(scExp, random_iter = 5, reduction = "PCA",
    corr_threshold = 99, percent_correlation = 1,
    downsample = 2500, verbose = TRUE, n_process = 250,
    BPPARAM = BiocParallel::bpparam()){
    if(ncol(scExp) < 5000) scExp@metadata$Unfiltered <- scExp
    pca_t = Matrix::t(SingleCellExperiment::reducedDim(scExp, reduction))
    correlation_values <- vector(length = random_iter)
    corChIP <- as.matrix(SingleCellExperiment::reducedDim(scExp, "Cor"))
    gc()
    limitC <- 0
    downsample = min(downsample, ncol(pca_t)) 
    if(verbose) message("ChromSCape::filter_correlated_cell_scExp - ",
    "Calculating correlation threshold using random matrix values...")
    for (i in seq_len(random_iter)) {
        random_mat <- matrix(
            sample(pca_t[,sample(seq_len(ncol(pca_t)),
                                 size = downsample)]), nrow = dim(pca_t)[1])
        threshold <-
            quantile(coop::pcor(x = random_mat, inplace = TRUE), probs = seq(0, 1, 0.01))
        limitC <- threshold[corr_threshold + 1]
        correlation_values[i] = limitC
    }
    limitC_mean = mean(correlation_values, na.rm = TRUE)
    rm(pca_t, random_mat)
    gc()
    if(verbose) message("ChromSCape::filter_correlated_cell_scExp - ",
                        "Filtering low correlated cells...")
    
    system.time({
    selection_cor_filtered =  unlist(DelayedArray::blockApply(
        corChIP, grid = DelayedArray::colAutoGrid(
            corChIP, ncol = min(ncol(corChIP), n_process)),
        BPPARAM = BPPARAM, verbose = FALSE, 
        function(X) apply(X, 2, function(x) length(which(x > limitC_mean)))
    ))
    })
    gc()
    selection_cor_filtered = selection_cor_filtered > 
        (percent_correlation * 0.01) * nrow(corChIP)
    
    scExp <- scExp[, selection_cor_filtered]
    gc()
    SingleCellExperiment::reducedDim(scExp, "Cor") = 
        as(SingleCellExperiment::reducedDim(scExp, "Cor")[,selection_cor_filtered], "dspMatrix")
    gc()

    if(verbose) message("ChromSCape::filter_correlated_cell_scExp - ",
                        "Re-calculating hierarchical clustering...")
    d = stats::as.dist(1 - as.matrix(SingleCellExperiment::reducedDim(scExp,"Cor")))
    gc()
    hc_cor_cor_filtered <- stats::hclust(d, method = "ward.D")
    gc()
    hc_cor_cor_filtered$labels = rep("", ncol(scExp))
    scExp@metadata$hc_cor = hc_cor_cor_filtered
    gc()
    
    for(alt in SingleCellExperiment::altExpNames(scExp)){
        SingleCellExperiment::reducedDim(SingleCellExperiment::altExp(scExp, alt), "Cor") = 
            as(SingleCellExperiment::reducedDim(SingleCellExperiment::altExp(scExp, alt), "Cor")[,selection_cor_filtered], "dspMatrix")
        d = stats::as.dist(1 - as.matrix(SingleCellExperiment::reducedDim(SingleCellExperiment::altExp(scExp, alt),"Cor")))
        gc()
        hc_cor_cor_filtered <- stats::hclust(d, method = "ward.D")
        gc()
        hc_cor_cor_filtered$labels = rep("", ncol(SingleCellExperiment::altExp(scExp, alt)))
        SingleCellExperiment::altExp(scExp, alt)@metadata$hc_cor = hc_cor_cor_filtered
        SingleCellExperiment::altExp(scExp, alt)@metadata$limitC = limitC_mean
    }
    gc()
    
    scExp@metadata$limitC = limitC_mean #specific to filtered scExp
    return(scExp)
}

```


