---
title: "global_var"
author: "Anna Schwager"
date: "07/08/2024"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Directory setup
The mainDir is the root directory. All the other paths are relative to mainDir.
```{r directories creation}
mainDir = here::here()
inputDir =  file.path(mainDir,"input") 
outputDir = file.path(mainDir, "output")
annotDir = file.path(mainDir, "annotation")

```
                                                                                               
## Loading libraries      

```{r libraries_new, message=F, warning=F}

library(Signac)
library(Seurat)
library(EnsDb.Mmusculus.v79)
library(EnsDb.Hsapiens.v86)
library(tidyverse)
library(ChIPpeakAnno)
library(biovizBase)
library(ggsci)
library(scales)
library(umap)
library(pheatmap)
library(RColorBrewer)
library(ggpointdensity)
library(grid)
library(gridExtra)
library(ggvenn)

```

## Loading annotation
```{r annotation, message=F, warning=F}
hg38 = seqlengths(EnsDb.Hsapiens.v86)
names(hg38) = paste0("chr", names(hg38))
# extract gene annotations from EnsDb
annotations_hg38 <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
# change to UCSC style
seqlevels(annotations_hg38) <- paste0('chr', seqlevels((annotations_hg38)))
genome(annotations_hg38) <- "hg38"

mm10 = seqlengths(EnsDb.Mmusculus.v79)
names(mm10) = paste0("chr", names(mm10))
# extract gene annotations from EnsDb
annotations_mm10 <- GetGRangesFromEnsDb(ensdb = (EnsDb.Mmusculus.v79))
# change to UCSC style 
seqlevels(annotations_mm10) <- paste0('chr', seqlevels(annotations_mm10))
genome(annotations_mm10) <- "mm10"

```

## Plot themes
Some explanation 
```{r colors}
mypal <- pal_npg("nrc", alpha = 0.9)(10)
mypal <- c(mypal, "#996699", "#e0d0e0")

#scales::show_col(mypal)

#c("#D3D3D3", "#BECFCC", "#A9C6BE", "#55B3A4", "#00A087E5", "#005043", "#00201b")

umap_blues = c("#D3D3D3", "#A3C5D0", "#4DBBD5E5", "#006D7F", "#00201b")
umap_greens = c("#D3D3D3", "#BECFCC", "#00A087E5", "#005043", "#00201b")

```

```{r session_info}
sessionInfo()
```

R version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: Europe/Paris
tzcode source: internal

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] ggvenn_0.1.10              gridExtra_2.3              ggpointdensity_0.1.0       RColorBrewer_1.1-3         pheatmap_1.0.12           
 [6] umap_0.2.10.0              scales_1.3.0               ggsci_3.2.0                biovizBase_1.54.0          ChIPpeakAnno_3.40.0       
[11] lubridate_1.9.4            forcats_1.0.0              stringr_1.5.1              dplyr_1.1.4                purrr_1.0.2               
[16] readr_2.1.5                tidyr_1.3.1                tibble_3.2.1               ggplot2_3.5.1              tidyverse_2.0.0           
[21] EnsDb.Hsapiens.v86_2.99.0  EnsDb.Mmusculus.v79_2.99.0 ensembldb_2.30.0           AnnotationFilter_1.30.0    GenomicFeatures_1.58.0    
[26] AnnotationDbi_1.68.0       Biobase_2.66.0             GenomicRanges_1.58.0       GenomeInfoDb_1.42.1        IRanges_2.40.1            
[31] S4Vectors_0.44.0           BiocGenerics_0.52.0        Seurat_5.2.0               SeuratObject_5.0.2         sp_2.1-4                  
[36] Signac_1.14.0             

loaded via a namespace (and not attached):
  [1] ProtGenerics_1.38.0         matrixStats_1.5.0           spatstat.sparse_3.1-0       bitops_1.0-9                httr_1.4.7                 
  [6] InteractionSet_1.34.0       backports_1.5.0             tools_4.4.2                 sctransform_0.4.1           R6_2.5.1                   
 [11] lazyeval_0.2.2              uwot_0.2.2                  withr_3.0.2                 prettyunits_1.2.0           VennDiagram_1.7.3          
 [16] progressr_0.15.1            textshaping_0.4.1           cli_3.6.3                   formatR_1.14                spatstat.explore_3.3-4     
 [21] fastDummies_1.7.4           labeling_0.4.3              spatstat.data_3.1-4         askpass_1.2.1               ggridges_0.5.6             
 [26] pbapply_1.7-2               systemfonts_1.2.0           Rsamtools_2.22.0            foreign_0.8-88              dichromat_2.0-0.1          
 [31] parallelly_1.41.0           limma_3.62.2                BSgenome_1.74.0             rstudioapi_0.17.1           RSQLite_2.3.9              
 [36] generics_0.1.3              BiocIO_1.16.0               ica_1.0-3                   spatstat.random_3.3-2       Matrix_1.7-1               
 [41] futile.logger_1.4.3         abind_1.4-8                 lifecycle_1.0.4             yaml_2.3.10                 SummarizedExperiment_1.36.0
 [46] SparseArray_1.6.0           BiocFileCache_2.14.0        Rtsne_0.17                  blob_1.2.4                  promises_1.3.2             
 [51] crayon_1.5.3                pwalign_1.2.0               miniUI_0.1.1.1              lattice_0.22-6              cowplot_1.1.3              
 [56] KEGGREST_1.46.0             pillar_1.10.1               knitr_1.49                  rjson_0.2.23                future.apply_1.11.3        
 [61] codetools_0.2-20            fastmatch_1.1-6             glue_1.8.0                  spatstat.univar_3.1-1       data.table_1.16.4          
 [66] vctrs_0.6.5                 png_0.1-8                   spam_2.11-0                 gtable_0.3.6                cachem_1.1.0               
 [71] xfun_0.50                   S4Arrays_1.6.0              mime_0.12                   survival_3.8-3              RcppRoll_0.3.1             
 [76] statmod_1.5.0               fitdistrplus_1.2-2          ROCR_1.0-11                 nlme_3.1-166                bit64_4.6.0-1              
 [81] progress_1.2.3              filelock_1.0.3              RcppAnnoy_0.0.22            rprojroot_2.0.4             irlba_2.3.5.1              
 [86] rpart_4.1.24                KernSmooth_2.23-26          Hmisc_5.2-2                 colorspace_2.1-1            DBI_1.2.3                  
 [91] nnet_7.3-20                 tidyselect_1.2.1            bit_4.5.0.1                 compiler_4.4.2              curl_6.1.0                 
 [96] httr2_1.0.7                 graph_1.84.1                htmlTable_2.4.3             xml2_1.3.6                  DelayedArray_0.32.0        
[101] plotly_4.10.4               rtracklayer_1.66.0          checkmate_2.3.2             lmtest_0.9-40               RBGL_1.82.0                
[106] rappdirs_0.3.3              digest_0.6.37               goftest_1.2-3               spatstat.utils_3.1-2        rmarkdown_2.29             
[111] XVector_0.46.0              base64enc_0.1-3             htmltools_0.5.8.1           pkgconfig_2.0.3             MatrixGenerics_1.18.1      
[116] dbplyr_2.5.0                regioneR_1.38.0             fastmap_1.2.0               rlang_1.1.4                 htmlwidgets_1.6.4          
[121] UCSC.utils_1.2.0            shiny_1.10.0                farver_2.1.2                zoo_1.8-12                  jsonlite_1.8.9             
[126] BiocParallel_1.40.0         VariantAnnotation_1.52.0    RCurl_1.98-1.16             magrittr_2.0.3              Formula_1.2-5              
[131] GenomeInfoDbData_1.2.13     dotCall64_1.2               patchwork_1.3.0             munsell_0.5.1               Rcpp_1.0.14                
[136] reticulate_1.40.0           stringi_1.8.4               zlibbioc_1.52.0             MASS_7.3-64                 plyr_1.8.9                 
[141] parallel_4.4.2              listenv_0.9.1               ggrepel_0.9.6               deldir_2.0-4                Biostrings_2.74.1          
[146] splines_4.4.2               tensor_1.5                  multtest_2.62.0             hms_1.1.3                   igraph_2.1.2               
[151] spatstat.geom_3.3-4         RcppHNSW_0.6.0              reshape2_1.4.4              biomaRt_2.62.0              futile.options_1.0.1       
[156] XML_3.99-0.18               evaluate_1.0.3              universalmotif_1.24.2       lambda.r_1.2.4              tzdb_0.4.0                 
[161] httpuv_1.6.15               openssl_2.3.1               RANN_2.6.2                  polyclip_1.10-7             future_1.34.0              
[166] scattermore_1.2             xtable_1.8-4                restfulr_0.0.15             RSpectra_0.16-2             later_1.4.1                
[171] ragg_1.3.3                  viridisLite_0.4.2           memoise_2.0.1               GenomicAlignments_1.42.0    cluster_2.1.8              
[176] timechange_0.3.0            globals_0.16.3              here_1.0.1 

